---
title: "Class 8 Exercises"
output: html_document
date: "2024-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
```

## Exercises

### Easy

Do all the easy exercises from Chapter 9: **9E1** through **9E6**.

### Medium

#### 9M1

Re-estimate the terrain ruggedness model from the chapter, but now using a uniform prior and
an exponential prior for the standard deviation, sigma. The uniform prior should be dunif(0,10)
and the exponential should be dexp(1). Do the different priors have any detectible influence on the
posterior distribution?

```{r code from the book with modification}

## R code 9.11
library(rethinking)
data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000)
dd <- d[ complete.cases(d$rgdppc_2000) , ]
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )

## R code 9.12
m8.3 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dunif(0,10) # dexp( 1 )
    ) , data=dd )

precis( m8.3 )

```


#### 9M2

The Cauchy and exponential priors from the terrain ruggedness model are very weak. They
can be made more informative by reducing their scale. Compare the dcauchy and dexp priors for
progressively smaller values of the scaling parameter. As these priors become stronger, how does each
influence the posterior distribution?

```{r}

m8.3_cauchy <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dcauchy( 0.1 )
    ) , data=dd )
# Your solution herre

m8.3_exp <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 0.001 )
    ) , data=dd )
# Your solution herre

precis( m8.3_cauchy, depth = 2)

precis( m8.3_exp, depth = 2)

```

### Hard

#### 9H1

Run the model below and then inspect the posterior distribution and explain what it is accomplishing.

```{r}

mp <- map2stan(alist(
  a ~ dnorm(0,1),
  b ~ dcauchy(0,1)
  ),
  data=list(y=1),
  start=list(a=0,b=0),
  iter=1e4, warmup=100 , WAIC=FALSE )

```

Compare the samples for the parameters a and b. Can you explain the different trace plots, using
what you know about the Cauchy distribution?

#### 9H2

```{r}

# Your solution herre

```
